---
title: "NYC Food Orders"
author: "Dereje Pollock, Kaylee Wu, Elijah Pilhe"
format: pdf
editor: visual
geometry: "left=1.5cm,right=1.5cm,top=2cm,bottom=2cm"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.height=4, fig.width=8)
```

```{r}
# load packages - add any you would like
library(tidyverse)
library(lme4)
library(lmerTest)
library(gridExtra)
library(knitr)
library(kableExtra)
install.packages("tinytex")
```

## Abstract

A 150 word abstract.

The number of restaurants in New York is steadily increasing, reflecting the city’s fast-paced lifestyle and the growing reliance on dining out or ordering in. With so many options available, understanding what factors influence the cost of a meal is essential for both restaurant efficiency and customer satisfaction. Our project aims to explore how specific restaurant characteristics contribute to the cost of an order.


\newpage

## Background and Significance

Type background and significance section here.

## Data


We are using a multilevel dataset titled "NYC Restaurants Data - Food Ordering and Delivery." This dataset contains variables such as order_id, restaurant, cuisine_type, WeekendOrWeekday, rating, delivery_time, and prep_time. The dataset is structured into two levels: level 2 variables include restaurant, WeekendOrWeekday, and cuisine type. While level 1 variables include rating, cost, prep_time, order_id, and delivery_time.


## **Key Variables Considered**

### **Response Variable**

-   **cost**: Total price of the order.

### **Predictors**

-   **prep_time**: Time taken to prepare the order (numeric).

-   **rating**: Customer rating of the restaurant (numeric).

-   **cuisine_type**: Type of cuisine (categorical: Italian, Chinese, etc.).

-   **WeekendOrWeekday**: Whether the order was placed on a weekday or weekend (binary).

-   **restaurant**: Random effect to account for differences across restaurants (factor)

Type data section here. Include captions with tables and graphs. Examples:

```{r}
FoodOrdersNYC <- read.csv("FoodOrdersNYC.csv")


```

```{r, fig.cap="First 10 Observations in our dataset"}
table1 <- FoodOrdersNYC %>%
  filter(row_number() < 11 )
kable(table1, booktabs=T, caption="The First 10 observations") %>%
  kable_styling(font_size = 8, latex_options = c("scale_down", "striped"))
```
# Prep Time vs Cost

We can use a scatterplot to understand the relationship between the time it takes to prepare an order, and the cost of the order.

```{r fig.cap="Cost vs. Prep Time"}
# Improved scatter plot with a fitted regression line
p_enhanced <- ggplot(FoodOrdersNYC, aes(x = prep_time, y = cost)) +
  geom_point(alpha = 0.6, color = "blue", size = 1.5) +  # Adjusted point color, size, and transparency
  geom_smooth(method = "lm", color = "red", se = TRUE) +  # Add a linear regression line with confidence interval
  labs(
    x = "Prep Time (minutes)",
    y = "Cost"
  ) +
  theme(
    plot.subtitle = element_text(face = "italic"),  # Italic subtitle
    
  )

# Print the enhanced plot
print(p_enhanced)
```

we can see a positive correlation between prep time and cost of an order. This correlation does not apear to be dramatic, but if we separate orders by their cuisine type, we can hopefully see clearer trends.




# Prep time vs Cost by Cuisine
```{r, fig.cap="Cost vs. Prep Time by Cuisine"}

# Create the scatter plot with adjustments for better display
p_total_time_by_cuisine <- ggplot(FoodOrdersNYC, aes(x = prep_time, y = cost)) +
  geom_point(alpha = 0.6, color = "blue") +  # Add points
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add fitted line
  facet_wrap(~cuisine_type, scales = "free", nrow = 2) +  # Facet by cuisine with two rows
  theme_minimal() +  # Minimal theme
  theme(
    axis.text.x = element_text(size = 8),  # Smaller text size for x-axis labels
    axis.text.y = element_text(size = 8),  # Smaller text size for y-axis labels
    strip.text.x = element_text(size = 10, angle = 0)  # Smaller facet labels, no angle
  ) +
  labs(
    x = "Prep Time",
    y = "Cost ($)"
  )

# Print the plot
print(p_total_time_by_cuisine)
```

When we separate the orders by Cuisine type, we can see that the effect of prep time on cost varies beteen cuisine types. This indicates that we should include a random slope in our model. We can see that American Cuisine has many more orders, and the number of restaurants that serve American Cuisine is also greater than some of our other Cuisine types. Japanese Cuisine also has considerably more orders.

```{r}
# Filter for American cuisine
american_restaurants <- FoodOrdersNYC[FoodOrdersNYC$cuisine_type == "American", ]

```


```{r}
# Costvs Prep Time for American Cuisines
p_american_revised <- ggplot(american_restaurants, aes(x = prep_time, y = cost)) +
  geom_point(alpha = 0.6, color = "blue", size = 2) +  # Adjust point size
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add fitted line
  facet_wrap(~restaurant, scales = "free", ncol = 5) +  # Organize facets
  scale_x_continuous(limits = c(15, 40), breaks = seq(15, 40, 5)) +  # Adjust x-axis limits and breaks
  theme_minimal(base_size = 8) +  # Increase base font size for better readability
  theme(
    strip.text.x = element_text(size = 9, angle = 0, hjust = 0.5),  # Adjust facet label size and angle
    panel.grid.major.x = element_line(color = "gray80"),  # Add major x grid lines
    panel.grid.minor.x = element_line(color = "gray90", linetype = "dotted")  # Add minor x grid lines
  ) +
  labs(
    title = "Cost vs Prep Time for American Cuisine by Restaurant",
    x = "Preparation Time (minutes)",
    y = "Cost ($)",
    subtitle = "Each panel represents a different American cuisine restaurant"
  )

# Print the plot
print(p_american_revised)
```



```{r}
# Filter for Japanese cuisine
Japanese_restaurants <- FoodOrdersNYC[FoodOrdersNYC$cuisine_type == "Japanese", ]
```

```{r}
# Revised plot with adjusted axis limits and improved aesthetics
p_japanese <- ggplot(Japanese_restaurants, aes(x = prep_time, y = cost)) +
  geom_point(alpha = 0.6, color = "blue", size = 2) +  # Adjust point size
  geom_smooth(method = "lm", se = FALSE, color = "red") +  # Add fitted line
  facet_wrap(~restaurant, scales = "free", ncol = 5) +  # Organize facets
  scale_x_continuous(limits = c(15, 40), breaks = seq(15, 40, 5)) +  # Adjust x-axis limits and breaks
  theme_minimal(base_size = 8) +  # Increase base font size for better readability
  theme(
    strip.text.x = element_text(size = 9, angle = 0, hjust = 0.5),  # Adjust facet label size and angle
    panel.grid.major.x = element_line(color = "gray80"),  # Add major x grid lines
    panel.grid.minor.x = element_line(color = "gray90", linetype = "dotted")  # Add minor x grid lines
  ) +
  labs(
    title = "Cost vs Prep Time for Japanese Cuisine by Restaurant",
    x = "Preparation Time (minutes)",
    y = "Cost ($)",
    subtitle = "Each panel represents a different Japanese cuisine restaurant"
  )

# Print the plot
print(p_japanese)
```








## Methods


# M1


Consider $Y_{ij}$ to be the cost of the $j$-th order, in the $i$-th restaurant.


$$
Y_{ij} = \beta_0 + \alpha_1 \text{cuisineType}_i + \beta_2 \text{prepTime}_{ij} + u_i + \epsilon_ij
$$


$$
\quad \text{where} \quad u_i \sim \mathcal{N}(0, \sigma_u^2) \quad \text{and} \quad \epsilon_{ij} \sim \mathcal{N}(0, \sigma^2)
$$

```{r, echo=TRUE, include=FALSE}
# Simplified model with only a random intercept for each restaurant
m1 <- lmer(cost ~ prep_time +  (1 | restaurant), data = FoodOrdersNYC)

summary(m1)
```

```{r, echo=TRUE}
# 
# Linear mixed model fit by REML. t-tests use Satterthwaite's method ['lmerModLmerTest']
# Formula: cost ~ prep_time + (1 | restaurant)
#    Data: FoodOrdersNYC
# 
# REML criterion at convergence: 6036.9
# 
# Scaled residuals: 
#     Min      1Q  Median      3Q     Max 
# -1.8081 -0.7616 -0.2614  0.7832  2.1464 
# 
# Random effects:
#  Groups     Name        Variance Std.Dev.
#  restaurant (Intercept)  0.4086  0.6392  
#  Residual               58.9305  7.6766  
# Number of obs: 872, groups:  restaurant, 30
# 
# Fixed effects:
#              Estimate Std. Error        df t value Pr(>|t|)    
# (Intercept)  16.26660    1.57969 821.48717  10.297   <2e-16 ***
# prep_time     0.07546    0.05654 866.93661   1.335    0.182    
# ---
# Signif. codes:  0 ‘***’ 0.001 ‘**’ 0.01 ‘*’ 0.05 ‘.’ 0.1 ‘ ’ 1
# 
# Correlation of Fixed Effects:
#           (Intr)
# prep_time -0.982


```

```{r}
# Create diagnostic plots for the model without log transformation
# 1. Residuals vs. Fitted Values
p1_no_log <- ggplot(NULL, aes(x = fitted(m1), y = residuals(m1))) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(x = "Fitted Values", y = "Residuals", title = "Residuals vs. Fitted Values (No Log)") +
  theme_minimal()

# 2. Normal Q-Q Plot
p2_no_log <- ggplot(NULL, aes(sample = residuals(m1))) +
  stat_qq() +
  stat_qq_line() +
  labs(title = "Normal Q-Q Plot (No Log)") +
  theme_minimal()

# 3. Histogram of Residuals
p3_no_log <- ggplot(NULL, aes(x = residuals(m1))) +
  geom_histogram(bins = 30, fill = "lightblue", color = "black") +
  labs(x = "Residuals", title = "Histogram of Residuals (No Log)") +
  theme_minimal()


# Arrange the plots for the model without log transformation
grid.arrange(p1_no_log, p2_no_log, p3_no_log, nrow = 2, ncol = 2)

```






Type methods section here. Write your final model in equation form, like this:

```{=tex}
\begin{align*}
Y_{ij} & = [\alpha_{0}+\alpha_{1}\textrm{Orch}_{i}+\beta_{0}\textrm{LargeEns}_{ij}+\beta_{1}\textrm{Orch}_{i}\textrm{LargeEns}_{ij}] \\
    & + [u_{i}+v_{i}\textrm{LargeEns}_{ij}+\epsilon_{ij}]
\end{align*}
```
where,

$$
\left[ \begin{array}{c}
            u_{i} \\ v_{i}
          \end{array}  \right] \sim N \left( \left[
          \begin{array}{c}
            0 \\ 0
          \end{array} \right], \left[
          \begin{array}{cc}
            \sigma_{u}^{2} & \rho_{uv}\sigma_{u}\sigma_v \\
            \rho_{uv}\sigma_{u}\sigma_v & \sigma_{v}^{2}
          \end{array} \right] \right) 
$$

and $\epsilon_{ij}\sim N(0,\sigma^2)$.

## Results

Write your results section here. Include the fixed and random effects tables, as below.


From the model output, most variables are not statistically significant. However, the p-value for food preparation time (0.057) is very close to the conventional significance threshold, suggesting a potential relationship between preparation time and cost. Specifically, for every additional minute of preparation time, the cost is expected to increase by 0.072, assuming all other factors remain constant. While this result is not definitively significant at the 0.05 level, it indicates a possible trend that longer preparation times may contribute to higher order costs. More analysis will be added.

```{r}

```

## Discussion and Conclusions

Type discussion section.



# Conclusion

1.  what does the result indicate about the restaurants in NYC
2.  What customers and owners can learn
3.  The limitations of the model

\newpage







## References

References in APA format. For example:

Sadler, M. E., & Miller, C. J. (2010). Performance anxiety: A longitudinal study of the roles of personality and experience in musicians. Social Psychological and Personality Science, 1(3), 280-287.

Lin, M. C. (2019). An Investigation Of Music Performance Anxiety In Taiwanese Pianists, Vocalists, String And Wind Instrumentalists At The College Level.

Stoeber, J., & Eismann, U. (2007). Perfectionism in young musicians: Relations with motivation, effort, achievement, and distress. Personality and Individual Differences, 43(8), 2182-2192.

Roback, P., & Legler, J. (2021). Beyond Multiple Linear Regression: Applied Generalized Linear Models And Multilevel Models in R. CRC Press.

\newpage

## Appendix (optional)
